Introduction to R and RStudio

### 2018-08-29 DATA606/Lab0/Michael-intro_to_r.Rmd
Additional work

---
title: "Introduction to R and RStudio"
output:
  html_document:
    css: ./lab.css
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

@@

![rinterface](more/rInterface2014.png)

@@

```{r load-abrbuthnot-data, eval=TRUE}
source("more/arbuthnot.R")
```

@@

```{r try-it-1}
arbuthnot$girls
```
Notice that the way R has printed these data is different. When we looked at the
complete data frame, we saw 82 rows, one on each line of the display. These data
are no longer structured in a table with other variables, so they are displayed 
one right after another. Objects that print out in this way are called *vectors*;
they represent a set of numbers. R has added numbers in [brackets] along the 
left side of the printout to indicate locations within the vector. For example,
`5218` follows `[1]`, indicating that `5218` is the first entry in the vector. 
And if `[43]` starts a line, then that would mean the first number on that line
would represent the 43<sup>rd</sup> entry in the vector.

@@

3.  Now, make a plot of the proportion of boys over time. What do you see? Tip: 
If you use the up and down arrow keys, you can scroll through your previous 
commands, your so-called command history. You can also access it by clicking on 
the history tab in the upper right panel.  This will save you a lot of typing in
the future.
```{r try-it-3}
share_boys <- arbuthnot$boys / (arbuthnot$boys + arbuthnot$girls)
plot(x = arbuthnot$year, y = share_boys, main = 'More boys than girls (share > 50%)')
```

@@

```{r load-present-data, eval=TRUE}
source("more/present.R")
```

@@

@@

@@

@@

>>>>

---
title: "Introduction to R and RStudio"
output:
  html_document:
    css: ./lab.css
    highlight: pygments
    theme: cerulean
    keep_md: yes
  pdf_document: default
---

@@

![ ](more/rInterface2014.png)

@@

```{r load-abrbuthnot-data, eval=TRUE}
source("~/Documents/CUNY-SPS/DATA606/Lab0/more/arbuthnot.r")
```

@@

```{r try-it-1}
arbuthnot$girls
```

Notice that the way R has printed these data is different. When we looked at the
complete data frame, we saw 82 rows, one on each line of the display. These data
are no longer structured in a table with other variables, so they are displayed 
one right after another. Objects that print out in this way are called *vectors*;
they represent a set of numbers. R has added numbers in [brackets] along the 
left side of the printout to indicate locations within the vector. For example,
`5218` follows `[1]`, indicating that `5218` is the first entry in the vector. 
And if `[43]` starts a line, then that would mean the first number on that line
would represent the 43<sup>rd</sup> entry in the vector.

@@

3.  Now, make a plot of the proportion of boys over time. What do you see? Tip: 
If you use the up and down arrow keys, you can scroll through your previous 
commands, your so-called command history. You can also access it by clicking on 
the history tab in the upper right panel.  This will save you a lot of typing in
the future.

```{r try-it-3}
share_boys <- arbuthnot$boys / (arbuthnot$boys + arbuthnot$girls)
plot(x = arbuthnot$year, y = share_boys, main = 'More boys than girls (share > 50%)')
```

@@

```{r load-present-data, eval=TRUE}
source("~/Documents/CUNY-SPS/DATA606/Lab0/more/present.R")
```

@@

```{r}
unique(present$year)
```

**1940 to 2002**

```{r}
dim(present)
```

**63 rows and 3 columns**

```{r}
names(present)
```

@@

```{r warning=FALSE}
suppressMessages(library(dplyr))

combined <- rbind(arbuthnot, present) %>%
  mutate(total = boys + girls) %>%
  mutate(Dataset = ifelse(year > 1939, 'Present', 'Arbuthnot'))

combined %>%
  group_by(Dataset) %>%
  summarise(`Average records per year` = mean(total))
```

**Present dataset has significantly more records per year than Arbuthnot's**

@@

```{r}
library(ggplot2)
combined %>%
  mutate(`Boy-to-Girl Ratio` = boys / girls) %>%
  ggplot(., aes(x = Dataset, y = `Boy-to-Girl Ratio`, fill = Dataset)) + 
  geom_boxplot(show.legend = FALSE) 
```

**Arbuthnot's observations does not hold in the Present dataset**

@@

```{r}
combined[combined$year == max(combined$year),]$year
```
